### 2022-04-29 paper/main/manuscript.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                                 %%
%% Please do not use \input{...} to include other tex files.       %%
%% Submit your LaTeX manuscript as one .tex document.              %%
%%                                                                 %%
%% All additional figures and files should be attached             %%
%% separately and not embedded in the \TeX\ document itself.       %%
%%                                                                 %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%\documentclass[referee,sn-basic]{sn-jnl}% referee option is meant for double line spacing

%%=======================================================%%
%% to print line numbers in the margin use lineno option %%
%%=======================================================%%

%%\documentclass[lineno,sn-basic]{sn-jnl}% Basic Springer Nature Reference Style/Chemistry Reference Style

%%======================================================%%
%% to compile with pdflatex/xelatex use pdflatex option %%
%%======================================================%%

%%\documentclass[pdflatex,sn-basic]{sn-jnl}% Basic Springer Nature Reference Style/Chemistry Reference Style

%%\documentclass[sn-basic]{sn-jnl}% Basic Springer Nature Reference Style/Chemistry Reference Style
% \documentclass[sn-mathphys]{sn-jnl}% Math and Physical Sciences Reference Style
%%\documentclass[sn-aps]{sn-jnl}% American Physical Society (APS) Reference Style
%%\documentclass[sn-vancouver]{sn-jnl}% Vancouver Reference Style
%%\documentclass[sn-apa]{sn-jnl}% APA Reference Style
%%\documentclass[sn-chicago]{sn-jnl}% Chicago-based Humanities Reference Style
% \documentclass[sn-standardnature]{sn-jnl}% Standard Nature Portfolio Reference Style
% \documentclass[default]{sn-jnl}% Default
\documentclass[default,iicol,lineno]{sn-jnl}% Default with double column layout

%%%% Standard Packages
%%<additional latex packages if required can be included here>
%%%%

%%%%%=============================================================================%%%%
%%%%  Remarks: This template is provided to aid authors with the preparation
%%%%  of original research articles intended for submission to journals published
%%%%  by Springer Nature. The guidance has been prepared in partnership with
%%%%  production teams to conform to Springer Nature technical requirements.
%%%%  Editorial and presentation requirements differ among journal portfolios and
%%%%  research disciplines. You may find sections in this template are irrelevant
%%%%  to your work and are empowered to omit any such section if allowed by the
%%%%  journal you intend to submit to. The submission guidelines and policies
%%%%  of the journal take precedence. A detailed User Manual is available in the
%%%%  template package for technical guidance.
%%%%%=============================================================================%%%%

\jyear{2021}%

%% as per the requirement new theorem styles can be included as shown below
\theoremstyle{thmstyleone}%
\newtheorem{theorem}{Theorem}%  meant for continuous numbers
%%\newtheorem{theorem}{Theorem}[section]% meant for sectionwise numbers
%% optional argument [theorem] produces theorem numbering sequence instead of independent numbers for Proposition
\newtheorem{proposition}[theorem]{Proposition}%
%%\newtheorem{proposition}{Proposition}% to get separate numbers for theorem and proposition etc.

\theoremstyle{thmstyletwo}%
\newtheorem{example}{Example}%
\newtheorem{remark}{Remark}%

\theoremstyle{thmstylethree}%
\newtheorem{definition}{Definition}%

\raggedbottom
%%\unnumbered% uncomment this for unnumbered level heads
%! path of figs

\graphicspath{{/Users/songshgeo/Documents/Pycharm/WGRegimes_YRB_2020/figures/}}
\begin{document}

\title[Article Title]{Identifying regime transitions for water governance at a basin scale}

%%=============================================================%%
%% Prefix	-> \pfx{Dr}
%% GivenName	-> \fnm{Joergen W.}
%% Particle	-> \spfx{van der} -> surname prefix
%% FamilyName	-> \sur{Ploeg}
%% Suffix	-> \sfx{IV}
%% NatureName	-> \tanm{Poet Laureate} -> Title after name
%% Degrees	-> \dgr{MSc, PhD}
%% \author*[1,2]{\pfx{Dr} \fnm{Joergen W.} \spfx{van der} \sur{Ploeg} \sfx{IV} \tanm{Poet Laureate}
%%                 \dgr{MSc, PhD}}\email{iauthor@gmail.com}
%%=============================================================%%

\author[1,2]{\fnm{Shuang} \sur{Song}}\email{songshgeo@mail.bnu.edu.cn}

\author[1,2]{\fnm{Shuai} \sur{Wang}}\email{shuaiwang@bnu.edu.cn}
% \equalcont{These authors contributed equally to this work.}

\author[1,2]{\fnm{Xutong} \sur{Wu}}\email{wuxutong@bnu.edu.cn}
% \equalcont{These authors contributed equally to this work.}
\author[3]{\fnm{Yongping} \sur{Wei}}\email{yongping.wei@uq.edu.au}
\author[4]{\fnm{Graeme S.} \sur{Cumming}}\email{graeme.cumming@jcu.edu.au}
\author*[1,2]{\fnm{Bojie} \sur{Fu}}\email{fubj@bnu.edu.cn}

\affil*[1]{\orgdiv{State Key Laboratory of Earth Surface Processes and Resource Ecology}, \orgname{Beijing Normal University,}, \orgaddress{\street{Xinjiekouwai Street}, \city{Beijing}, \postcode{100875}, \state{Beijing}, \country{China}}}

\affil[2]{\orgdiv{nstitute of Land Surface System and Sustainability}, \orgname{Beijing Normal University}, \orgaddress{\street{Xinjiekouwai Street}, \city{Beijing}, \postcode{100875}, \state{Beijing}, \country{China}}}

\affil[3]{\orgdiv{School of Earth and Environmental Sciences}, \orgname{The University of Queensland}, \orgaddress{\city{Brisbane}, \postcode{4067}, \state{QLD}, \country{Australia}}}

\affil[4]{\orgdiv{ARC Centre of Excellence for Coral Reef Studies}, \orgname{James Cook University}, \orgaddress{\city{Townsville}, \postcode{4811}, \state{QLD}, \country{Australia}}}

%%==================================%%
%% sample for unstructured abstract %%
%%==================================%%

\abstract{
    In many large river basins, water governance regimes from natural to social-ecological or `hydrosocial' situations profoundly change who gets water, when, and how. Therefore, identifying regime changes is critical to understanding the transition and guiding sustainability in water governance. We combined three main dimensions -stress, purpose and allocation- to develop a quantitative Integrated Water Governance Index (IWGI) to detect water governance regimes at a basin scale. Applying the IWGI to a rapidly-changing basin (the Yellow River Basin, China) describes how water governance shifts between massive supply, transformation governance, and adaptation-oriented regimes. In the YRB, the regime shifts' underlying causes were increasing water supply and demand before the governance transformation and re-allocation and regulation after the transformation. The IWGI offers a comprehensive and straightforward approach to linking the water governance regimes to the following challenges in sustainability, along with the hydrosocial transition.
}

\keywords{
    Regime shifts,
    Water governance,
    Transformation,
    Social-hydrology,
    Sustainability
}

%%\pacs[JEL Classification]{D8, H51}
%%\pacs[MSC Classification]{35A01, 65L10, 65L12, 65L20, 65L70}

\maketitle

\section{Introduction}\label{sec1}
\input{01_introduction}

\section{Results}\label{sec2}
\input{03_results}

\section{Discussion}\label{sec12}
\input{04_discussion}

\section{Conclusion}\label{sec13}
\input{05_conclusion}

\section{Methods}\label{sec11}
\input{02_methods}

\backmatter

\bmhead{Supplementary information}
When calculating the indicators (especially the SFV water stress index and the allocation entropy metric), we should use data at a lower (regional scale in this study) spatial scale. Therefore, we divide the YRB into four regions: source region (SR), upper region (UR), middle region (MR), and lower region (LR), according to characteristics and customary practices in the \ref{secA1}.
The formulation in detail for applying the SFV-index is available in the \ref{secA2}.
We used multiple sources of datasets in this study, \textit{Appendix}~\ref{secA3} introduces where they came from and how we harmonise them for analysis.

\bmhead{Acknowledgments}

Funding was provided by the National Natural Science Foundation of China (CN) (Grant Nos. NSFC 42041007).

\section*{Declarations}

% Some journals require declarations to be submitted in a standardised format. Please check the Instructions for Authors of the journal to which you are submitting to see if you need to complete this section. If yes, your manuscript must contain the following sections under the heading `Declarations':

\begin{itemize}
\item Funding
\item Conflict of interest/Competing interests (check journal-specific guidelines for which heading to use)
\item Ethics approval
\item Consent to participate
\item Consent for publication
\item Availability of data and materials
\item Code availability
\item Authors' contributions
\end{itemize}

\noindent
If any of the sections are not relevant to your manuscript, please include the heading and write `Not applicable' for that section.

%%===================================================%%
%% For presentation purpose, we have included        %%
%% \bigskip command. please ignore this.             %%
%%===================================================%%
\bigskip
\begin{flushleft}%
Editorial Policies for:

\bigskip\noindent
Springer journals and proceedings: \url{https://www.springer.com/gp/editorial-policies}

\bigskip\noindent
Nature Portfolio journals: \url{https://www.nature.com/nature-research/editorial-policies}

\bigskip\noindent
\textit{Scientific Reports}: \url{https://www.nature.com/srep/journal-policies/editorial-policies}

\bigskip\noindent
BMC journals: \url{https://www.biomedcentral.com/getpublished/editorial-policies}
\end{flushleft}

\begin{appendices}

\section{YRB Regions}\label{secA1}
\input{../sup/01_yrb.tex}

\section{SFV-index}\label{secA2}
\input{../sup/02_indicator.tex}

% \begin{figure}[tb!]
% 	\centering
% 	\includegraphics[width=0.9\linewidth]{sup/S3_wci.pdf}
% 	\caption{
% 	    Ratio of irrigated areas with water-saving equipments.
% 	}
% 	\label{fig:wci}
% \end{figure}

\begin{figure}[tb]
	\centering
	\includegraphics[width=0.6\linewidth]{sup/reservoirs.pdf}
	\caption{
	    Numbers of new reservoirs in each year.
	}
	\label{fig:reservoirs}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{sup/sensitivity.pdf}
    \caption{
        Sensitivity analysis of the threshold of p-values.
        \textbf{A.} number of breakpoints in different p-values, the scheme with two-breakpoints are the dominant situation.
        \textbf{B.} Threshold of p-values $\alpha=0.0005$.
        \textbf{C.} Threshold of p-values $\alpha=0.05$.
    }
    \label{fig:sensitivity}
\end{figure}

\section{Datasets}\label{secA3}
\input{../sup/03_datasets.tex}

\end{appendices}

%%===========================================================================================%%
%% If you are submitting to one of the Nature Portfolio journals, using the eJP submission   %%
%% system, please include the references within the manuscript file itself. You may do this  %%
%% by copying the reference list from your .bbl file, paste it into the main manuscript .tex %%
%% file, and delete the associated \verb+\bibliography+ commands.                            %%
%%===========================================================================================%%

\bibliographystyle{sn-standardnature}
\bibliography{mybib}% common bib file
%% if required, the content of .bbl file can be included here once bbl is generated
%%\input sn-article.bbl

%% Default %%
%%\input sn-sample-bib.tex%

\end{document}