### 2022-07-12 paper/sup/supplementary_information.tex
feat: appendix separated version

\documentclass[default, sn-standardnature]{sn-jnl}% Default with double column layout

%%%% Standard Packages
%%<additional latex packages if required can be included here>
%%%%

%%%%%=============================================================================%%%%
%%%%  Remarks: This template is provided to aid authors with the preparation
%%%%  of original research articles intended for submission to journals published
%%%%  by Springer Nature. The guidance has been prepared in partnership with
%%%%  production teams to conform to Springer Nature technical requirements.
%%%%  Editorial and presentation requirements differ among journal portfolios and
%%%%  research disciplines. You may find sections in this template are irrelevant
%%%%  to your work and are empowered to omit any such section if allowed by the
%%%%  journal you intend to submit to. The submission guidelines and policies
%%%%  of the journal take precedence. A detailed User Manual is available in the
%%%%  template package for technical guidance.
%%%%%=============================================================================%%%%

\jyear{2022}%
\newgeometry{left=2cm, right=2cm}
\linespread{1.5}

%% as per the requirement new theorem styles can be included as shown below
\theoremstyle{thmstyleone}%
\newtheorem{theorem}{Theorem}%  meant for continuous numbers
%%\newtheorem{theorem}{Theorem}[section]% meant for sectionwise numbers
%% optional argument [theorem] produces theorem numbering sequence instead of independent numbers for Proposition
\newtheorem{proposition}[theorem]{Proposition}%
%%\newtheorem{proposition}{Proposition}% to get separate numbers for theorem and proposition etc.

\theoremstyle{thmstyletwo}%
\newtheorem{example}{Example}%
\newtheorem{remark}{Remark}%

\theoremstyle{thmstylethree}%
\newtheorem{definition}{Definition}%
% \usepackage{tikz}
% \newcommand*{\circled}[1]{\lower.7ex\hbox{\tikz\draw (0pt, 0pt)%
%     circle (.5em) node {\makebox[1em][c]{\small #1}};}}
\raggedbottom
%%\unnumbered% uncomment this for unnumbered level heads
%! path of figs

\graphicspath{{/Users/songshgeo/Documents/Pycharm/WGRegimes_YRB_2020/figures/}}
\begin{document}

\title{Supporting Materials of ``Identifying regime transitions for water governance at a basin scale''}

%%=============================================================%%
%% Prefix	-> \pfx{Dr}
%% GivenName	-> \fnm{Joergen W.}
%% Particle	-> \spfx{van der} -> surname prefix
%% FamilyName	-> \sur{Ploeg}
%% Suffix	-> \sfx{IV}
%% NatureName	-> \tanm{Poet Laureate} -> Title after name
%% Degrees	-> \dgr{MSc, PhD}
%% \author*[1,2]{\pfx{Dr} \fnm{Joergen W.} \spfx{van der} \sur{Ploeg} \sfx{IV} \tanm{Poet Laureate}
%%                 \dgr{MSc, PhD}}\email{iauthor@gmail.com}
%%=============================================================%%

\author[1]{\fnm{Shuang} \sur{Song}}\email{songshgeo@mail.bnu.edu.cn}

\author[1]{\fnm{Shuai} \sur{Wang}}\email{shuaiwang@bnu.edu.cn}
% \equalcont{These authors contributed equally to this work.}

\author[1]{\fnm{Xutong} \sur{Wu}}\email{wuxutong@bnu.edu.cn}
% \equalcont{These authors contributed equally to this work.}
\author[2]{\fnm{Yongping} \sur{Wei}}\email{yongping.wei@uq.edu.au}
\author[3]{\fnm{Graeme S.} \sur{Cumming}}\email{graeme.cumming@jcu.edu.au}
\author[4]{\fnm{Yue} \sur{Qin}}\email{qinyue@pku.edu.cn}
\author[5]{\fnm{Xilin} \sur{Wu}}\email{wuxilin20@mails.ucas.ac.cn}
\author*[1,5]{\fnm{Bojie} \sur{Fu}}\email{bfu@rcees.ac.cn}

\affil*[1]{\orgdiv{State Key Laboratory of Earth Surface Processes and Resource Ecology}, \orgname{Beijing Normal University,}, \orgaddress{\street{Xinjiekouwai Street}, \city{Beijing}, \postcode{100875}, \state{Beijing}, \country{China}}}

\affil[2]{\orgdiv{School of Earth and Environmental Sciences}, \orgname{The University of Queensland}, \orgaddress{\city{Brisbane}, \postcode{4067}, \state{QLD}, \country{Australia}}}

\affil[3]{\orgdiv{ARC Centre of Excellence for Coral Reef Studies}, \orgname{James Cook University}, \orgaddress{\city{Townsville}, \postcode{4811}, \state{QLD}, \country{Australia}}}

\affil[4]{\orgdiv{College of Environmental Sciences and Engineering} \orgname{Peking University}, \orgaddress{\city{Beijing}, \postcode{100875}, \state{Beijing}, \country{China}}}

\affil[5]{\orgdiv{State Key Laboratory of Urban and Regional Ecology, Research Center for Eco-Environmental Sciences}, \orgname{Chinese Academy of Sciences}, \orgaddress{\city{Beijing}, \postcode{100875}, \state{Beijing}, \country{China}}}
%%==================================%%
%% sample for unstructured abstract %%
%%==================================%%

%% Comment out or remove this line before generating final copy for submission; this will also remove the warning re: "Consecutive odd pages found".
% \instructionspage

\maketitle

%% Adds the main heading for the SI text. Comment out this line if you do not have any supporting information text

\section{YRB Regions}\label{secA1}
\input{01_yrb.tex}

\section{SFV-index}\label{secA2}
\input{02_indicator.tex}

% \begin{figure}[tb!]
% 	\centering
% 	\includegraphics[width=0.9\linewidth]{sup/S3_wci.pdf}
% 	\caption{
% 	    Ratio of irrigated areas with water-saving equipments.
% 	}
% 	\label{fig:wci}
% \end{figure}

\begin{figure}[tb]
    \centering
    \includegraphics[width=0.6\linewidth]{sup/reservoirs.pdf}
    \caption{
        Numbers of new reservoirs in each year.
    }
    \label{fig:reservoirs}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{sup/sensitivity.pdf}
    \caption{
        Sensitivity analysis of the threshold of p-values.
        \textbf{A.} number of breakpoints in different p-values, the scheme with two-breakpoints are the dominant situation.
        \textbf{B.} Threshold of p-values $\alpha=0.0005$.
        \textbf{C.} Threshold of p-values $\alpha=0.05$.
    }
    \label{fig:sensitivity}
\end{figure}

\section{Datasets}\label{secA3}
\input{03_datasets.tex}

% \dataset{dataset_one.txt}{Type or paste legend here.}

% \dataset{dataset_two.txt}{Type or paste legend here. Adding longer text to show what happens, to decide on alignment and/or indentations for multi-line or paragraph captions.}

\bibliographystyle{sn-standardnature}
\bibliography{mybib}% common bib file

\end{document}