### 2023-03-30 paper/letter/response_letter.tex
feat(paper): :memo: 准备返修论文

\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{lipsum} % to generate some filler text
\usepackage{fullpage}
\usepackage{mathrsfs}
\usepackage{svg}
\usepackage{xr}
\usepackage{amsthm,cite,url,graphicx,booktabs,lipsum,color,bm,caption,subcaption,soul}
\usepackage{pifont,tikz,paralist,multirow,amssymb}
\usepackage{xifthen}
\usepackage{enumerate}
\usepackage{titlesec}
\newcounter{reviewer}
\setcounter{reviewer}{0}
\newcounter{point}[reviewer]
\setcounter{point}{0}

% This refines the format of how the reviewer/point reference will appear.
%\renewcommand{\thepoint}{C\,\thereviewer.\arabic{point}}

\renewcommand{\thepoint}{Comment\,\thereviewer.\arabic{point}}

% command declarations for reviewer points and our responses

\newenvironment{point1}
   {\refstepcounter{point} \bigskip \noindent {\textbf{Associate Editor Comment~}} ---\ }
   {\par }

\newcommand{\reviewersection}{\stepcounter{reviewer} \bigskip \hrule
                  \section*{Reviewer \thereviewer}}

% \newenvironment{point}

%    {\refstepcounter{point} \bigskip \noindent {\textbf{Reviewer~\thepoint} } ---\ }
%    {\par }

\newenvironment{point}
   {\refstepcounter{point} \bigskip \noindent {\textbf{Reviewer~\thepoint} } ---\ }
   {\par }

\newcommand{\shortpoint}[1]{\refstepcounter{point}   \bigskip \noindent
	{\textbf{Reviewer~Point~\thepoint} } ---~#1  }

\newenvironment{reply}
   {\medskip \noindent \color{blue} \begin{sf}\textbf{Reply}:\  }
   {\medskip \par \end{sf}}

\newcommand{\shortreply}[2][]{\medskip \noindent \begin{sf}\textbf{Reply}:\  #2
	\ifthenelse{\equal{#1}{}}{}{ \hfill \footnotesize (#1)}%
	\medskip \end{sf}}

\makeatletter
\newcommand*{\addFileDependency}[1]{% argument=file name and extension
  \typeout{(#1)}
  \@addtofilelist{#1}
  \IfFileExists{#1}{}{\typeout{No file #1.}}
}
\makeatother

\newcommand*{\myexternaldocument}[1]{%
    \externaldocument{#1}%
    \addFileDependency{#1.tex}%
    \addFileDependency{#1.aux}%
}

\myexternaldocument{main}
%\externaldocument[I-]{main}

\def\thesubsubsection{\arabic{subsubsection}}
\titleformat{\subsubsection}[runin]
{\normalfont\bfseries}
{\indent\thesubsubsection)}{0.5em}{}

\begin{document}
\section*{Notes on revision made to manuscript Paper $\#$ABC-2099-12-1234
}
\vspace{10pt}

The authors would like to thank the editor and reviewers for their constructive comments and suggestions that have helped improve the quality of this manuscript. The manuscript has undergone a thorough revision according to the editor and reviewers’ comments. Please see below our responses. For the reviewers’ convenience, we have highlighted significant changes in the revised manuscript in \textcolor{blue}{blue}.

% Let's start point-by-point with Reviewer 1

\section*{Response to the Associate Editor}
% General intro text goes here
\begin{point1}

    Dear Authors, thanks for the submission of the manuscript. We have asked two experts in the relevant fields to review the manuscript. Both reviewers overall find the merit in the manuscript however they also noted several major issues in the current manuscript. With my own reading, I fully agree with the reviewers and would like to highlight a few major issues:

    1. Overall the results and discussion are relatively shallow and I would encourage the authors to do deep dive into the results, e.g. change points, as also pointed out by both reviewers,

    2. I fully agree with Reviewer \#2 that more thorough review of existing and relevant indicators with IWGI, and

    3. I would encourage the authors to look into more recent data (beyond 2013) as pointed out by Reviewer \#1. In addition, I find the data in supplementary materials is not accessible, which should be amended. The data access and reproduction of the methodology/results are very important for WRR and our scientific community, so we take the issue related with data accessibility very seriously. Overall, based on the reviewers' recommendations, I would like to ask the authors to submit a suitably revised manuscript in due course.
\end{point1}

\begin{reply}
\input{response_v6/associate_editor.tex}

\end{reply}

\section*{Response to the reviewers}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\reviewersection
% General intro text goes here

% Point one description

%%%%%%%%%%%%%%%%
\begin{point}
	In the new submission, the paper should be read and corrected by a native English speaker as there are several verbal and syntactic errors. Also, the references should be checked and provided correctly. \label{comment_11}
\end{point}

% Our reply
\begin{reply}
	Thank you so much for your constructive comments. We have thus scrutinized the manuscript and corrected the grammar errors and typos.  Also, we have also corrected the format errors in the references.
\end{reply}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\reviewersection

%%%%%%%%%%%%%%%%
\begin{point}
\label{comment_21}a
\end{point}

\begin{reply}
 Thank you so much for your constructive comments.
\end{reply}

%%%%%%%%%%%%%%%%
\begin{point}
\label{comment_22}
\end{point}

\begin{reply}
 Thank you so much for your constructive comments.
\end{reply}
%%%%%%%%%%%%%%%%
\begin{point}

\end{point}

\begin{reply}
 Thank you so much for your constructive comments.
\end{reply}
%%%%%%%%%%%%%%%%
\begin{point}
\label{comment_31}
\end{point}

\begin{reply}
 Thank you so much for your constructive comments.
\end{reply}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\reviewersection

%%%%%%%%%%%%%%%%
\begin{point}

\end{point}

\begin{reply}
 Thank you so much for your constructive comments.
\end{reply}

%\bibliographystyle{IEEEtran}
%\bibliography{IEEEabrv,mybibfile}
\end{document}

\begin{point}

\end{point}

\begin{reply}

\end{reply}