### 2023-05-15 paper/letter/cover_letter_r1.tex
docs(paper): :memo: 更新cover letter和给编辑的回复信

\documentclass[11pt,a4paper,roman]{moderncv}

\graphicspath{{../../../figs/}}
\linespread{1.5}
\usepackage[english]{babel}
\usepackage{ragged2e}
\usepackage{float}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}

\moderncvstyle{classic}
\moderncvcolor{blue} % Bullet point color

% Page margins
\usepackage[scale=0.85]{geometry} % Page margins

% Your Information, please revise
\name{Shuai}{Wang}
\address{Haidian District, Beijing, China}{No.19, Xinjiekouwai St.}
\phone[mobile]{+86 182 1009 3639}
\email{shuaiwang@bnu.edu.cn}

\begin{document}

% Insert Olin Logo
\begin{minipage}[t]{\textwidth}
\includegraphics[width=0.40\textwidth]{bnu}
\end{minipage}

\recipient{Dear Editor in chief,}{}
\opening{\vspace*{-2em}}
\closing{Sincerely,}{\vspace*{-2em}}
\enclosure[Enclosures]{Main manuscript, Appendix}
\makelettertitle
% https://pythonjishu.com/qggkmtfnuj
% 调整两端对齐
\justify

Thank you for the opportunity to revise our manuscript titled ``Identifying regime transitions for water governance at a basin scale'' We appreciate the constructive feedback provided by the reviewers and believe that our revised manuscript more effectively addresses their concerns.

% 很抱歉我们花了这么长时间来返修，但正如您看到的，我们在新的版本中补充了很多内容和图表，努力回应审稿人和编辑的一切关切。我们也花时间试着联系了数据的作者团队。在得知正文中数据短时间内无法增补到最新年份后，我们使用其它数据源再次计算了IWGI，从而努力增强结果的鲁棒性。
I'm sorry it took us so long to go back, but as you can see, we've added a lot of content and graphics to the new version in an effort to respond to all the concerns of reviewers and editors. We also took the time to try to contact the team that authored the data. After learning that the data in the text could not be updated to the latest year for a short time, we supplement analysis using other data sources in an effort to enhance the robustness of the results.
In response to the reviewers' feedback, we also restructured the introduction and discussion sections to provide a more supportive and solid content of the results.

We believe that these revisions have significantly improved the quality and value of our manuscript, and we hope that you will find it suitable for publication in Water Resources Research. We are confident that our study makes a valuable contribution to the understanding of water governance in China and beyond.

The revised manuscript includes the main body of text with $4423$ words and $4$ figures; the supporting information with $4$ sections has $9$ supporting figures and $1$ supporting tables. Furthermore, we declare no conflicts of interest to disclose.

\vspace{0.3cm}

Thanks for your time. We appreciate your consideration!

\vspace{0.3cm}

Sincerely,

\textbf{Shuai Wang} (On behalf of the author team)

%\begin{minipage}[t]{\textwidth}
%	\includegraphics[height=1cm]{signature.jpeg}
%\end{minipage}

State Key Laboratory of Earth Surface Processes and Resource Ecology,

Faculty of Geographical Science,

Beijing Normal University,

Beijing 100875, China

Email: bfu@rcees.ac.cn

\end{document}