### 2023-10-06 paper/response_letter/source/editor.tex
docs(paper): :memo: prepare revise the paper

% chktex-file 46
\section*{Associate Editor}\label{editor}

\subsection{Remarks to Author}
\RC{} Dear Authors, thank you for the carful revisions. Our original referees have re-evaluated the revised manuscripts. While they are satisfied with most of the revisions, they find that some of their earlier comments were not fully addressed.

\RC{} As Reviewer \#1 pointed out, there needs to be more carful evaluation of the method with water stress calculation, which would affect part of the outcome.

\AR{}

\RC{} Reviewer \#2 provided a set of very useful comments to further improve the clarity of the manuscript. Reviewer \#2 also pointed out that the link in the SI is not accessible so please carefully check. We won't be able to pursue publication without warranting all the supporting information and data access. I would like to encourage the authors to address all the comments and re-submit a suitably revised manuscript in due course.

\AR{} Thank you for your constructive comments and for providing us with the opportunity to address the concerns raised by both reviewers. We appreciate the valuable feedback, which has helped us identify areas for improvement in our manuscript. We have carefully considered each point and have made substantial revisions to address these issues. Please find our detailed responses to the highlighted concerns below.

\AR*{} We hope that these revisions address the concerns raised by both reviewers. We are confident that these changes have significantly improved the quality of our manuscript and made it more relevant and valuable to the scientific community. We look forward to your further feedback and the possibility of our manuscript being accepted for publication.