### 2023-10-06 paper/response_letter/source/reviewer_2.tex

\section{Reviewer \#2}\label{reviewer_2}

Recommendation (Required): Return to author for minor revisions
Significant: The paper has some unclear or incomplete reasoning but will likely be a significant contribution with revision and clarification.
Supported: Yes
Referencing: Yes
Quality: Yes, it is well-written, logically organized, and the figures and tables are appropriate.
Data: Yes
Accurate Key Points: Yes

\subsection*{General comments}

Thank you for sharing your revised manuscript. The revised manuscript was very carefully constructed. This revision solved most of my questions, but there were some incomplete and vague parts in the response to "2.7. Comment \#7".

I agree that the change of the system is still a result of the accumulation of water governance practices in period. However, it was clear from your results and discussion that these system changes are dictated by specific factors (historical events): e.g. the Chinese slogan "human will conquer nature," rapid expansion of irrigated farmland and water diversion facilities so on, in the P1. I suggest that you organize them in a table that shows those factors (historical events) that are causing the "emergence" of each complex system. (Figure 4 is too conceptual for the historical phenomenon you were describing.) I believe that this will make the characteristics of each regime clearer.

The "social atmosphere" is an interesting term, but very vague. First of all, please define what it is (I have never heard of such a term in previous studies in my field). At the very least, you should discuss what "social atmosphere" is in the Discussion. Furthermore, I really wonder how your research and the proposed index can be practically contribute to understanding "the successes and failures of water governanceis" created by the "social atmosphere". Perhaps this "social atmosphere" is the most troubling and important factor to understand in water governance regime.

\subsection*{Minor comment}

I couldn't access the link http://www.yrcc.gov.cn/hhyl/hhjs/ in the Supporting Informations. Please confirm it.