### 2023-10-10 paper/letter/cover_letter_r2.tex

\documentclass[11pt,a4paper,roman]{moderncv}

\graphicspath{{../../../figs/}}
\linespread{1.5}
\usepackage[english]{babel}
\usepackage{ragged2e}
\usepackage{float}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}

\moderncvstyle{classic}
\moderncvcolor{blue} % Bullet point color

% Page margins
\usepackage[scale=0.85]{geometry} % Page margins

% Your Information, please revise
\name{Shuai}{Wang}
\address{Haidian District, Beijing, China}{No.19, Xinjiekouwai St.}
\phone[mobile]{+86 182 1009 3639}
\email{shuaiwang@bnu.edu.cn}

\begin{document}

% Insert Olin Logo
\begin{minipage}[t]{\textwidth}
\includegraphics[width=0.40\textwidth]{bnu}
\end{minipage}

\recipient{Dear Editor in chief,}{}
\opening{\vspace*{-2em}}
\closing{Sincerely,}{\vspace*{-2em}}
\enclosure[Enclosures]{Main manuscript, Appendix}
\makelettertitle
% https://pythonjishu.com/qggkmtfnuj
% 调整两端对齐
\justify

% 我们写信是为了提交我们的手稿的修订版，题为“确定中国黄河流域水治理的政权过渡”，供您和审稿人考虑。我们非常感谢审稿人提供的建设性意见，这些意见有助于提高我们工作的质量和清晰度。
We are writing to submit the revised version of our manuscript titled ``Identifying regime transitions for water governance at the Yellow River Basin, China'' for your and the reviewers’ consideration. We are grateful for the constructive feedback provided by the reviewers, which has been instrumental in enhancing the quality and clarity of our work.

One of the significant revisions in our manuscript involves the robustness check for the identified breakpoints in our study. Furthermore, we have addressed all the specific comments and suggestions raised by the reviewers meticulously. We introduced a new table and expanded the discussion to provide more comprehensive insights. These clarifications and modifications made in response to both reviewers’ inputs, have significantly enriched the content and narrative of our manuscript.

We believe that these revisions have addressed the reviewers’ concerns effectively and hope that the enhancements made will meet their expectations and standards of quality for publication in \textit{Water Resources Research}.

The revised manuscript includes the main body of text with $4423$ words and $4$ figures with $1$ table; the supporting information with $4$ sections has $9$ supporting figures and $2$ supporting tables. Furthermore, we declare no conflicts of interest to disclose.

\vspace{0.3cm}

Thanks for your time. We appreciate your consideration!

\vspace{0.3cm}

Sincerely,

\textbf{Shuai Wang} (On behalf of the author team)

%\begin{minipage}[t]{\textwidth}
%	\includegraphics[height=1cm]{signature.jpeg}
%\end{minipage}

State Key Laboratory of Earth Surface Processes and Resource Ecology,

Faculty of Geographical Science,

Beijing Normal University,

Beijing 100875, China

Email: shuaiwang@bnu.edu.cn

\end{document}