### 2023-10-10 paper/main/tab_regimes.tex
fix(paper): :bug: fixing paper response letter problems
based on Shuai Wang's comment

% Table generated by Excel2LaTeX from sheet 'Sheet1'
\begin{table}[htbp]
    \centering
    \caption{Brief summary of regime transitions for water governance at the Yellow River Basin}
      \begin{tabularx}{\textwidth}{p{0.15\textwidth} p{0.15\textwidth} p{0.2\textwidth} p{0.5\textwidth}}
      \toprule
      Time  & Closet social atmosphere$^*$ & Regime & Main trait \\
      \midrule
      $1965-1978$ & 1960s: Conquering nature & massive supply regime & boost water supply for services (mainly provisioning purposes then -livestock and crops) by constructing reservoirs and channels \\
      $1979-2001$ & 1978: reform and opening-up & governance transforming regime & triggered institutional changes for rapid enhancement of overall social adaptive capacity to severe water stresses \\
      $2002-2013$ & 2002: Environmental regulating & adaptation-oriented regime & adapting to stable high water stress with trade-offs between water-dependent regions and sectors \\
      \bottomrule
      \end{tabularx}\label{tab:regimes}%

      \footnotesize
      $^*$ Here, social atmosphere refers to the sociocultural context in which people live or in which something happens, including the culture that the individual was educated or lives.
  \end{table}%