### 2023-10-10 paper/sup/tab_robustness.tex

\begin{table}[h!]
    \centering
    \begin{tabularx}{\textwidth}{llll}
        \toprule
        \textbf{Method} & \textbf{Change points} & \textbf{Parameters$^*$} & \textbf{Reference} \\
        \midrule
        Generalized ESD Test & 1975, 1978, 2001 & No & \cite{matteson2014} \\
        Pettitt Method & 1978, 2001 & No & \cite{pettitt1979} \\
        Pelt & 1975, 2000 & $pen=0.05$ & \cite{killick2012} \\
        Binary Segmentation & 1975, 2000 & $bkp=2$ & \cite{bai1997} \\
        Bottom-Up Segmentation & 1975, 2000 & $bkp=2$ & \cite{keogh2001} \\
        \bottomrule
    \end{tabularx}
    \footnotesize{$pen$: Penalty coefficient; $bkp$: Pre-set number of change points. $No$: This method is a non-parametric test.}
    \caption{Comparison of different change points detection methods}\label{table:breakpoint-detection-methods}
\end{table}